---
title: "boostrap"
format: html
editor: visual
---

#Instructions

```{r}
#Create a table (in HTML or PDF) that shows the differences in median (with the BCa bootstrap 
#95% confidence intervals) maternal, infant, neonatal, and under-5 mortality between the
#countries exposed to versus not exposed to armed conflict for the year 2017. Be specific about
#the sample sizes used in each statistic. The table should be fully reproducible. Push the
#script that creates the table to your GitHub repository.
```

#Clear history

```{r}
#clear history
rm(list=ls(all=T)) 
```

#Load Libraries

```{r}
#load libraries
library(dplyr)
library(tidyr)
library(here)
library(boot)
library(tidyverse)
library(knitr)
library(flextable)
```

#Load datasets

```{r}
#load final data
data <- read.csv(here("clean", "final_data.csv"), header = TRUE)
data2017 <- data %>%
  filter(year == 2017) %>%
  drop_na(Maternal.Mortality.rate, Infant.mortality.rate, Under.5.mortality.rate, Neonatal.mortality.rate)

# Subset the Data Frame for the year 2017

matmor.arm1 <- data %>%
  filter(year==2017 & !is.na(Maternal.Mortality.rate) & armconf1 ==1) %>%
  select(country_name, Maternal.Mortality.rate)
matmor.arm0 <- data %>%
  filter(year==2017 & !is.na(Maternal.Mortality.rate) & armconf1 ==0) %>%
  select(country_name, Maternal.Mortality.rate)
```

#Perform bootstrap (traditional method)

```{r}
B <- 1000

med.diff <- rep(NA, B)
for(b in 1:B){
  resample.arm1 <- matmor.arm1[sample(nrow(matmor.arm1), size = nrow(matmor.arm1), replace = TRUE),]
  resample.arm0 <- matmor.arm0[sample(nrow(matmor.arm0), size = nrow(matmor.arm0), replace = TRUE),]
  med.diff[b] <- median(resample.arm1$Maternal.Mortality.rate) - median(resample.arm0$Maternal.Mortality.rate)
}
  
head(resample.arm1, 12)
```

##histogram of bootstrap distribution

```{r}
hist(med.diff, main = "Distribution of boostrap statistic")
```

#Bootstrap using Boot package ##maternal mortality

```{r}
getmeddiff <- function(data, indices) {
  sample_data <- data[indices, ]
  group_meds <- tapply(sample_data$Maternal.Mortality.rate, sample_data$armconf1, FUN = median)
  med.diff <- group_meds[2] - group_meds[1]
  return(med.diff)
}

bootout_mat <- boot(data2017, statistic = getmeddiff, strata = data2017$armconf1, R=1000)
bootout_mat
```

###bootstrap statistics

```{r}
#get confidence intervals
mat_ci <- boot.ci(boot.out = bootout_mat, conf = 0.95, type = "bca")

#bootstrap statistics for matmor: t1= 126.5, bias= 14.4095, SE= 64.77766, CI= 43.8, 290.3)
```

##Infant Mortality

```{r}
#bootstrap for infant mor
getmeddiff <- function(data, indices) {
  sample_data <- data[indices, ]
  group_meds <- tapply(sample_data$Infant.mortality.rate, sample_data$armconf1, FUN = median)
  med.diff <- group_meds[2] - group_meds[1]
  return(med.diff)
}

bootout_infant <- boot(data2017, statistic = getmeddiff, strata = data2017$armconf1, R=1000)
bootout_infant
```

###bootstrap statistics

```{r}
#get confidence intervals
infant_ci <- boot.ci(boot.out = bootout_infant, conf = 0.95, type = "bca")

#bootstrap statistics for infant mor: t1= 19.8, bias= -0.9521, SE= 5.943379, CI= 6.60, 29.16)
```

##Under 5 Mortality

```{r}
#bootstrap for under-5 mor
getmeddiff <- function(data, indices) {
  sample_data <- data[indices, ]
  group_meds <- tapply(sample_data$Under.5.mortality.rate, sample_data$armconf1, FUN = median)
  med.diff <- group_meds[2] - group_meds[1]
  return(med.diff)
}

bootout_under5 <- boot(data2017, statistic = getmeddiff, strata = data2017$armconf1, R=1000)
bootout_under5
```

###bootstrap statistics

```{r}
#get confidence intervals
under5_ci <- boot.ci(boot.out = bootout_under5, conf = 0.95, type = "bca")

#bootstrap statistics for Under mor: t1= 28.9, bias= -3.0935, SE= 9.678971, CI= 8.97, 45.68)
```

##Neonatal Mortality

```{r}
#bootstrap for neomor
getmeddiff <- function(data, indices) {
  sample_data <- data[indices, ]
  group_meds <- tapply(sample_data$Neonatal.mortality.rate, sample_data$armconf1, FUN = median)
  med.diff <- group_meds[2] - group_meds[1]
  return(med.diff)
}

bootout_neo <- boot(data2017, statistic = getmeddiff, strata = data2017$armconf1, R=1000)
bootout_neo
```

###boostrap statistics

```{r}
#get confidence intervals
neo_ci <- boot.ci(boot.out = bootout_neo, conf = 0.95, type = "bca")

#bootstrap statistics for neomor: t1= 12.4, bias= -0.8052, SE= 3.455128, CI= 5.4, 20.1)
```

#TABLE \## Create a data frame for each variable

```{r}
table0 <- data.frame(
  Mortality_Group = c('maternal', 'infant', 'neonatal', 'under5'),
  Median_Difference = c(bootout_mat$t0, bootout_infant$t0, bootout_under5$t0, bootout_neo$t0),
  Confidence_Interval_lower = c(mat_ci$bca[4], infant_ci$bca[4], under5_ci$bca[4], neo_ci$bca[4]),
  Confidence_Interval_upper = c(mat_ci$bca[5], infant_ci$bca[5], under5_ci$bca[5], neo_ci$bca[5]),
  n = c(sum(!is.na(data2017$Maternal.Mortality.rate)), sum(!is.na(data2017$Neonatal.mortality.rate)), sum(!is.na(data2017$Under.5.mortality.rate)), sum(!is.na(data2017$Infant.mortality.rate)))
)

table
```

##Add title
```{r}
table <- flextable(table0)
table <- add_header_row(
  x = table, values = c("Median Difference in Mortality between Countries exposed to vs not exposed to Armed Conflict in 2017"),
  colwidths = c(5))
table <- theme_box(table)

table
```
# Export as pdf and save it in "figures" folder in armed conflict folder

```{r}
#save as pdf
save_as_html(table, path = here("figures"), title = "mortality_ci_table")
```


